---
title: "Data manipulation"
output: github_document
---


```{r setup}
library(tidyverse)
library(dplyr)

```


##load in the FAS Litters Data

```{r }
litters_df = read_csv("./data/FAS_litters.csv",show_col_types = FALSE)
litters_df = janitor::clean_names(litters_df)

```


## `select`

#Choose some columns and not others.

select(dataset,column_name)

```{r}

select(litters_df,group,litter_number)
```


select a range of columns using `:`

```{r}
select(litters_df,group,gd0_weight:gd_of_birth)
```


Instead of saying what we want to keep, we can specify what we dont want using `-`

```{r}
select(litters_df, -litter_number)
```


Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```


Rename some columns without doing select

```{r}
rename(litters_df,GROUP = group, LITTer_NUmBer = litter_number)
```



#Select helpers : ?select_helpers

select everything that starts with "gd"

```{r}
select(litters_df, starts_with("gd"))
```


select litter_number as the first column and keep everything else 

```{r}
select(litters_df, litter_number, everything())
```


same as the above func, it relocates litter_number as the first column and keep the rest

```{r}
relocate(litters_df, litter_number)
```

`Select` is the key function, but there are funcs that can duplicate  to some of the things that you can do with select


## `filter`

`filter` is intended to remove rows or to keep rows that you are interested in retaining for some subsequent analysis

`select` operates on columns while `filter` operates on rows


Filter the data with the weight of mother on the first day that she's pregnant that is less than 22
```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

A list of logical operations is available on the P8105 website

us `==` as equal to

```{r}
filter(litters_df,gd_of_birth == 20)
```

Ways to negate

Get all of the data where gd_of_birth is not 20.
This works on more complicated things
```{r}
filter(litters_df,!(gd_of_birth == 20))
```

Another way of doing "not equal to"
```{r}
filter(litters_df,gd_of_birth != 20)
```


```{r}
filter(litters_df,gd0_weight >= 22,gd_of_birth == 20 )
```

```{r}
filter(litters_df, group == "Mod8")
```

filter where group is equal to either "Con7" or "Mod8"
```{r}
filter(litters_df, group %in% c("Con7","Mod8"))
```



## `mutate`

used to either modify an existing variable or create some new variables

options(tibble.width = Inf) # displays all columns.
options(tibble.print_max = Inf) # to show all the rows.

```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))      #making all group variables lower case         
```



## `arrange`

putting things in order


```{r}
arrange(litters_df, pups_born_alive)
```

first arrange in order by pups_born_alive and then within each of those levels its in order for gd0_weight

```{r}
arrange(litters_df, pups_born_alive,gd0_weight)
```



## `%>%`


```{r}
#reading dataset
litters_data_raw = read_csv(("./data/FAS_litters.csv"),show_col_types = FALSE)

#cleaning names in dataset
litters_clean_name = janitor::clean_names(litters_data_raw)

#select everything from the dataset except pups_survived
litters_data_selected = select(litters_clean_name, -pups_survive)

litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)

# drop everything where gd0_weight is missing / drop NAs
# use only litters_without_missing = drop_na(litters_mutated) to get rid of all NAs in data litters_mutated
litters_without_missing = drop_na(litters_mutated, gd0_weight)

```


USE THE PIPE OPERATOR INSTEAD!

use shift + command + m to type ` %>% `

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>%     # no need to put anything in this () because the pipes automatically generates the results from the first line into the ()
  select(-pups_survive) %>%           #select everything except pups_survive
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)

```

