---
title: "Data manipulation"
output: github_document
---


```{r setup}
library(tidyverse)

```


##load in the FAS Litters Data

```{r }
litters_df = read_csv("./data/FAS_litters.csv",show_col_types = FALSE)
litters_df = janitor::clean_names(litters_df)

```


## `select`

#Choose some columns and not others.

select(dataset,column_name)

```{r}

select(litters_df,group,litter_number)
```


select a range of columns using `:`

```{r}
select(litters_df,group,gd0_weight:gd_of_birth)
```


Instead of saying what we want to keep, we can specify what we dont want using `-`

```{r}
select(litters_df, -litter_number)
```


Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```


Rename some columns without doing select

```{r}
rename(litters_df,GROUP = group, LITTer_NUmBer = litter_number)
```



#Select helpers : ?select_helpers

select everything that starts with "gd"

```{r}
select(litters_df, starts_with("gd"))
```


select litter_number as the first column and keep everything else 

```{r}
select(litters_df, litter_number, everything())
```


same as the above func, it relocates litter_number as the first column and keep the rest

```{r}
relocate(litters_df, litter_number)
```

`Select` is the key function, but there are funcs that can duplicate  to some of the things that you can do with select


## `filter`

`filter` is intended to remove rows or to keep rows that you are interested in retaining for some subsequent analysis

`select` operates on columns while `filter` operates on rows


Filter the data with the weight of mother on the first day that she's pregnant that is less than 22
```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

A list of logical operations is available on the P8105 website

us `==` as equal to

```{r}
filter(litters_df,gd_of_birth == 20)
```

Ways to negate

Get all of the data where gd_of_birth is not 20.
This works on more complicated things
```{r}
filter(litters_df,!(gd_of_birth == 20))
```

Another way of doing "not equal to"
```{r}
filter(litters_df,gd_of_birth != 20)
```


```{r}
filter(litters_df,gd0_weight >= 22,gd_of_birth == 20 )
```

```{r}
filter(litters_df, group == "Mod8")
```

filter where group is equal to either "Con7" or "Mod8"
```{r}
filter(litters_df, group %in% c("Con7","Mod8"))
```


