---
title: "Data manipulation"
output: github_document
---


```{r setup}
library(tidyverse)

```


##load in the FAS Litters Data

```{r }
litters_df = read_csv("./data/FAS_litters.csv",show_col_types = FALSE)
litters_df = janitor::clean_names(litters_df)

```


## `select`

#Choose some columns and not others.

select(dataset,column_name)

```{r}

select(litters_df,group,litter_number)
```


select a range of columns using `:`

```{r}
select(litters_df,group,gd0_weight:gd_of_birth)
```


Instead of saying what we want to keep, we can specify what we dont want using `-`

```{r}
select(litters_df, -litter_number)
```


Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```


Rename some columns without doing select

```{r}
rename(litters_df,GROUP = group, LITTer_NUmBer = litter_number)
```



#Select helpers : ?select_helpers

select everything that starts with "gd"

```{r}
select(litters_df, starts_with("gd"))
```


select litter_number as the first column and keep everything else 

```{r}
select(litters_df, litter_number, everything())
```


same as the above func, it relocates litter_number as the first column and keep the rest

```{r}
relocate(litters_df, litter_number)
```

`Select` is the key function, but there are funcs that can duplicate  to some of the things that you can do with select




